Cell Decompositions for Rank Two Quiver Grassmannians

### 2019-05-11 response to reviewer.tex

\newcommand{\filt}[6]{
\[
\begin{xy}
\xymatrix@R20pt@C20pt{
&\mathbb{C}^3&\\\langle #1,#2 \rangle\ar[ru]&\langle #3,#4\rangle\ar[u]&\langle #5,#6\rangle\ar[lu]\\\langle #1\rangle\ar[u]&\langle #3\rangle\ar[u]&\langle #5\rangle\ar[u]}
\end{xy}
\]
} 
\begin{document}
\pagestyle{empty}

\noindent{\bf Response to the reviewer - Cell decompositions for rank two quiver Grassmannians}\\

\noindent In the following, we point out the changes we made.

Section 4:
\begin{enumerate}
\renewcommand{\labelenumi}{(\arabic{enumi})}
\item Note that the proof and the examples before show how to compute the kernel of the right hand vertical map for $m=2$ and $m=3$. As it is indeed not obvious to prove surjectivity for general $m$ directly, we deduce it from the $m=2,3$-cases and apply $\tau^{-1}$-shifts which preserve exact sequences if no direct summand is injective. This is mentioned in Remark 3.2 to which we refer to in the proof of Lemma 3.30.
\item We omit the notation $~\perp$ as it is only used in Corollary 3.32.
\item This should be a matter of definition and by reindexing appropriately we should get the version with $\ast$'s on the right.
\item We refer to Cerulli Irelli's paper now. But not that the assumptions in Theorem 1 are stronger than ours as we do not assume that every basis element is mapped on a scalar multiple of another basis element. Note also that the number of fixed points is finite in Cerulli Irelli's case which is not true in general.
\item We inserted a remark concerning the result of Haupt. Our approach seems to simplify the language and generalize the techniques of Haupt. If we are not mistaken, the proof of Theorem 1.2(c) in Section 8 only deals with those representations of the universal (abelian) covering whose dimension vectors only consists of $0$'s and $1$'s. We actually consider arbitrary representations of $\hat Q$. Moreover, the explicit description of the fixed points of $\mathrm{Gr}_{\mathbf{e}}(M)$ of a liftable representation $M$ as subrepresentations of the lifted representations $\hat M$ seems to be missing.

\item We inserted a remark which addresses this issue.
\item We inserted a remark in Theorem 4.22 (former 4.20) which is meant to explain how the statement is obtained from the claim. Moreover, it should make the proof clearer.
\item I find the strategy if 4.23 is very clear and I also think that the proof of 4.24 is very close to the suggestion of the referee. Any ideas what to change?
\end{enumerate}

### 2019-05-11 Grassmannian_cells.tex
further updates

The following orthogonality property is a primary reason we need to lift to the universal cover of $K(n)$.
\begin{corollary}
  \label{cor:perpendicular}
  Consider a non-empty subset $I\subsetneq\{1,\ldots,n\}$ and fix an element $j\in I$.
  For $m\geq 2$, we have $\tilde P_m(I,j)\in (\tilde P_{m+1}^I)^\perp$.
\end{corollary}
\begin{proof}
  If $m\geq 3$, we consider the long exact sequence obtained when applying $\Hom(\tilde P_{m+1}^I,-)$ to the sequence
  \[\ses{\tilde P_m(I,j)}{\tilde P_{m,j}}{\tau\tilde P_{m+1}^I}.\]
  Following Lemma~\ref{le:properties}.(3), we have $\Hom(\tilde P_{m+1}^I,\tilde P_{m,j})=0$ and using the long exact sequence this immediately implies $\Hom\big(\tilde P_{m+1}^I,\tilde P_m(I,j)\big)=0$.
  From Lemma~\ref{le:properties}.(3) again, we have $\Ext(\tilde P_{m+1}^I,\tilde P_{m,j})\cong\CC$.
  Using the Auslander-Reiten formulas \cite[Theorem IV.2.13]{ass} and Lemma~\ref{le:properties}.(4), we get
  \[\dim\Ext(\tilde P_{m+1}^I,\tau\tilde P_{m+1}^I)=\dim\Hom(\tilde P_{m+1}^I,\tilde P_{m+1}^I)=1.\]
  It follows that the surjective map $\Ext(\tilde P_{m+1}^I,\tilde P_{m,j})\to\Ext(\tilde P_{m+1}^I,\tau\tilde P_{m+1}^I)$ appearing in the long exact sequence is in fact an isomorphism.
  But the Auslander-Reiten formulas and Lemma~\ref{le:properties}.(4) again imply 
  \[\dim\Hom(\tilde P_{m+1}^I,\tau\tilde P_{m+1}^I)=\dim\Ext(\tilde P_{m+1}^I,\tilde P_{m+1}^I)=0.\]
  Combining with the preceding discussion, this gives $\Ext\!\big(\tilde P_{m+1}^I,\tilde P_m(I,j)\big)=0$.

@@

  \noindent{\bf Claim.}
  For any admissible sequence $\bfI=(I_0,\ldots,I_k)$ with $k<m$ and any subset $J\subset I_0$, the quiver Grassmannian $\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^J)$ admits a cell decomposition which is compatible with the unique map $\pi_{m+1}^{J,\bfI}:\tilde P_{m+1}^J\to\tilde P_{m+1}^\bfI$ from Lemma~\ref{le:truncated quotients} in the following sense:
  \begin{itemize}
    \item[($\dagger$)] For any $V\in\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^J)$ such that $\pi_{m+1}^{J,\bfI}(V)\neq 0$, we have $\pi_{m+1}^{J,\bfI}(V')\neq 0$ for all $V'\in C_V$, where $C_V$ is the affine cell which contains $V$.
  \end{itemize}
  \vs
Note that having proved the claim, for $|J|\leq n-2$ the statement itself is obtained when setting $\bfI=(J\cup\{i\})$ for any $i\in\{1,\ldots,n\}\backslash\{J\}$. For $|J|=n-1$, we can apply Lemma~\ref{le:truncated quotients}, i.e. $P_{m+1}^J\cong\tilde P_{m,j}^{\{j\}}$ with $J^c=\{j\}$, together with the fact that $\tilde P_{m,j}^{\{j\}}$ is obtained when translating $\tilde P_{m}^{\{j\}}$ in the universal covering quiver. Clearly, translation does not affect quiver Grassmannians when also translating dimension vectors $\tbfe$, i.e. quiver Grassmannians of $\tilde P_{m,j}^{\{j\}}$ are just the same as those of $\tilde P_{m}^{\{j\}}$.

Finally, recall that the preprojective representations $\tilde P_{m+1}$ and their quiver Grassmannians are obtained when setting $J=\varnothing$ and considering any short exact sequence
		$$\ses{\tilde P_{m,j}}{\tilde P_{m+1}}{\tilde P_{m+1}^{\{j\}}}$$
for $j\in\{1,\ldots,n\}$. Then we can use the fact that $\tilde P_{m,j}$ is the translation of $\tilde P_m$.

>>>>

The following orthogonality property is a primary reason we need to lift to the universal cover of $K(n)$.
\begin{corollary}
  \label{cor:perpendicular}
  Consider a non-empty subset $I\subsetneq\{1,\ldots,n\}$ and fix an element $j\in I$.
  For $m\geq 2$, we have $\Hom(\tilde P_{m+1}^I,\tilde P_m(I,j))=\Ext(\tilde P_{m+1}^I,\tilde P_m(I,j))=0$.
\end{corollary}
\begin{proof}
  If $m\geq 3$, we consider the long exact sequence obtained when applying $\Hom(\tilde P_{m+1}^I,-)$ to the sequence
  \[\ses{\tilde P_m(I,j)}{\tilde P_{m,j}}{\tau\tilde P_{m+1}^I}.\]
  Following Lemma~\ref{le:properties}.(3), we have $\Hom(\tilde P_{m+1}^I,\tilde P_{m,j})=0$ and using the long exact sequence this immediately implies $\Hom\big(\tilde P_{m+1}^I,\tilde P_m(I,j)\big)=0$.
  From Lemma~\ref{le:properties}.(3) again, we have $\Ext(\tilde P_{m+1}^I,\tilde P_{m,j})\cong\CC$.
  Using the Auslander-Reiten formulas \cite[Theorem IV.2.13]{ass} and Lemma~\ref{le:properties}.(4), we get
  \[\dim\Ext(\tilde P_{m+1}^I,\tau\tilde P_{m+1}^I)=\dim\Hom(\tilde P_{m+1}^I,\tilde P_{m+1}^I)=1.\]
  It follows that the surjective map $\Ext(\tilde P_{m+1}^I,\tilde P_{m,j})\to\Ext(\tilde P_{m+1}^I,\tau\tilde P_{m+1}^I)$ appearing in the long exact sequence is in fact an isomorphism.
  But the Auslander-Reiten formulas and Lemma~\ref{le:properties}.(4) again imply 
  \[\dim\Hom(\tilde P_{m+1}^I,\tau\tilde P_{m+1}^I)=\dim\Ext(\tilde P_{m+1}^I,\tilde P_{m+1}^I)=0.\]
  Combining with the preceding discussion, this gives $\Ext\!\big(\tilde P_{m+1}^I,\tilde P_m(I,j)\big)=0$.

@@

  \noindent{\bf Claim.}
  For any admissible sequence $\bfI=(I_0,\ldots,I_k)$ with $k<m$ and any subset $J\subset I_0$, the quiver Grassmannian $\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^J)$ admits a cell decomposition which is compatible with the unique map $\pi_{m+1}^{J,\bfI}:\tilde P_{m+1}^J\to\tilde P_{m+1}^\bfI$ from Lemma~\ref{le:truncated quotients} in the following sense:
  \begin{itemize}
    \item[($\dagger$)] For any $V\in\Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^J)$ such that $\pi_{m+1}^{J,\bfI}(V)\neq 0$, we have $\pi_{m+1}^{J,\bfI}(V')\neq 0$ for all $V'\in C_V$, where $C_V$ is the affine cell which contains $V$.
  \end{itemize}
  \vs

\begin{remark}
Note that having proved the claim, for $|J|\leq n-2$ the statement itself is obtained when setting $\bfI=(J\cup\{i\})$ for any $i\in\{1,\ldots,n\}\backslash\{J\}$. For $|J|=n-1$, we can apply Lemma~\ref{le:truncated quotients}, i.e. $P_{m+1}^J\cong\tilde P_{m,j}^{\{j\}}$ with $J^c=\{j\}$, together with the fact that $\tilde P_{m,j}^{\{j\}}$ is obtained when translating $\tilde P_{m}^{\{j\}}$ in the universal covering quiver. Clearly, translation does not affect quiver Grassmannians when also translating dimension vectors $\tbfe$, i.e. quiver Grassmannians of $\tilde P_{m,j}^{\{j\}}$ are just the same as those of $\tilde P_{m}^{\{j\}}$.

Furthermore, recall that the preprojective representations $\tilde P_{m+1}$ and their quiver Grassmannians are obtained when setting $J=\varnothing$ and considering any short exact sequence
		$$\ses{\tilde P_{m,j}}{\tilde P_{m+1}}{\tilde P_{m+1}^{\{j\}}}$$
for $j\in\{1,\ldots,n\}$. Then we can use the fact that $\tilde P_{m,j}$ is the translation of $\tilde P_m$.

Finally, note that the claim together with Proposition \ref{label} ensures that the fiber dimension along the points of the affine cell $C_V$ is constant. Thus Theorem \ref{vb} gives the statement of the Theorem.
\end{remark}
